---
  title: "Portfolio Returns"
---
  
```{r prior_code, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
packages <- c("tidyverse", "kableExtra", "knitr", "tidyquant")

for (package in packages) {
  if (!require(package, character.only=T, quietly=T)) {
    install.packages(package, repos="http://cran.us.r-project.org")
  }
  library(package, character.only=T)
}

options("scipen"=99)

options("digits"=3)

data_import <- function(x, na = "") {
  df <- read.csv(x, check.names = FALSE, na.strings = na)
  assign(substr(x, 6, nchar(x)-4), df, envir = .GlobalEnv)
  return(NULL)
}

filenames <- list.files("Data/", pattern = "*.csv", full.names = TRUE)

sapply(filenames, data_import, na = c("", "NA", "#N/A", "#N/A Invalid Security"))

price <- price[,colSums(is.na(price))<nrow(price)]

mcap <- mcap[, colnames(price)]

price[,-1] <- sapply(price[, -1],  as.numeric)
mcap[,-1] <- sapply(mcap[, -1], as.numeric)

price <- xts(price[,-1], order.by = ymd(price$Date))
mcap <- xts(mcap[,-1], order.by = ymd(mcap$Date))
jse <- xts(jse[,"JALSH Index"], order.by = ymd(jse[,1]))
rfr <- xts(rfr[,2], order.by = ymd(rfr$Date))

price <- price["2000/2019"]
mcap <- mcap["2000/2019"]
jse <- jse["2000/2019"]
rfr <- rfr["2000/2019"]

price <- price[endpoints(price)]
mcap <- mcap[index(price)]
jse <- jse[index(price)]
rfr <- rfr[index(price)] / 1200

returns <- diff(log(price))
jalsh <- diff(log(jse))

# Top 100 shares 

mcap100 <- t(apply(-mcap, 1, rank, na.last = "keep", ties.method = "first"))

mcap100_mask <- ifelse(mcap100 <= 100,1,NA)

mcap <- mcap * mcap100_mask

returns <- returns * mcap100_mask

vol <- rollapply(data = returns, width = 12, FUN = sd, fill = NA)
volAnn <- vol[endpoints(vol, on = "years")] * sqrt(12)

yrs <- year(index(volAnn))[-1]

for (yr in yrs) {
  mon <- as.character(yr)
  temp <- coredata(volAnn[mon])
  vol[mon, ] <- rbind(temp, temp[rep(1, 11), ])
}

vol100 <- t(apply(-vol, 1, rank, na.last = "keep", ties.method = "first"))

high_vol <- ifelse(vol100 <= 20, 1, NA)
vol_2 <- ifelse(vol100 > 20 & vol100 <= 40, 1, NA)
vol_3 <- ifelse(vol100 > 40 & vol100 <= 60, 1, NA)
vol_4 <- ifelse(vol100 > 60 & vol100 <= 80, 1, NA)
low_vol <- ifelse(vol100 > 80 & vol100 <= 100, 1, NA)

high_vol <- ifelse(vol100 <= quantile(vol100, 0.2, na.rm = TRUE), 1, NA)
vol_2 <- ifelse(vol100 > quantile(vol100, 0.2, na.rm = TRUE) & vol100 <= quantile(vol100, 0.4, 
                                                                                  na.rm = TRUE), 1, NA)
vol_3 <- ifelse(vol100 > quantile(vol100, 0.4, na.rm = TRUE) & vol100 <= quantile(vol100, 0.6, 
                                                                                  na.rm = TRUE), 1, NA)
vol_4 <- ifelse(vol100 > quantile(vol100, 0.6, na.rm = TRUE) & vol100 <= quantile(vol100, 0.8, 
                                                                                  na.rm = TRUE), 1, NA)
low_vol <- ifelse(vol100 > quantile(vol100, 0.8, 
                                    na.rm = TRUE), 1, NA)
```

Creating portfolio returns is easy once all the hard work is done. The code below demonstates how straightfoward this is.

```{r portfolio}
portfolio <- data.frame(high = rowMeans(returns * high_vol, na.rm = T),
                        port2 = rowMeans(returns * vol_2, na.rm = T),
                        port3 = rowMeans(returns * vol_3, na.rm = T),
                        port4 = rowMeans(returns * vol_4, na.rm = T),
                        low = rowMeans(returns * low_vol, na.rm = T),
                        row.names = index(returns))
```

The code above creates a dataframe containing returns of the five portfolios. We've combined a few steps together here so let's break them down. First, we get returns of the shares in each portfolio by multiplying the returns by the porfolio matrices. Since the portfolio matrices just have 1s and NAs, the result is returns where the share is in the portfolio, and NAs where it isn't.

After this, we get the average return in each month using rowMeans. Since the returns matrix at this stage will only have 20 non-missing data in each row, this average will be the return of the portfolio in that month. And just like that, we have a dataframe of returns for all portfolios.

```{r show_portfolios, results='hide'}
rbind(head(portfolio[-c(1:12),]), tail(portfolio))
```



```{r portfolio_table, echo=FALSE}
portfolios <- kable(rbind(head(portfolio[-c(1:12),]), tail(portfolio)), row.names = FALSE)
portfolios <- gsub("<table>", "<table class = 'pretty-table'>", portfolios)
portfolios <- gsub('> NA', ' class = "na-data"> NA', portfolios)
portfolios
remove(portfolios)
```

The final return calculation is to get the cumulative returns. This is done using the `cumprod` function.

```{r rolling_returns}
portfolio[1:12,] <- 0

r.xts <- xts(portfolio, order.by = index(returns))[-c(1:11),]
c.xts <- cumprod(1+r.xts)
port_ret <- data.frame(dates = index(-c.xts), c.xts)
```

The result is shown below

```{r show_returns, results='hide'}
rbind(head(port_ret[-c(1:12),]), tail(port_ret))
```



```{r returns_table, echo=FALSE}
port_retTable <- kable(rbind(head(port_ret[-c(1:12),]), tail(port_ret)), row.names = FALSE)
port_retTable <- gsub("<table>", "<table class = 'pretty-table'>", port_retTable)
port_retTable <- gsub('> NA', ' class = "na-data"> NA', port_retTable)
port_retTable
remove(port_retTable)
```

The output above does not give an intuitive understanding of each portfolio return. So we turn to `ggplot2`, the excellent package for graphics. The code below shows how to create a relatively simple cumulative return chart. 

```{r graph_results, dpi=600}
ggplot(port_ret, aes(x = dates, group = 1))+
  geom_line(aes(y = high, color = "High Vol"), size = 1)+
  geom_line(aes(y = low, color = "Low Vol"), size = 1)+
  scale_color_manual("", values = c("High Vol" = "#4da734", 
                                    "Low Vol" = "#3363fa"))+
  theme_minimal()+
  scale_x_date(date_breaks = "1 year", date_labels =  "%b %y")+
  theme(axis.text.x=element_text(angle=60, hjust=1),
        plot.background=element_rect(fill="#a5f3e4"),
        legend.position = "top",
        legend.margin = margin(t = -5, r = 0, b = -10, l = 0, unit = "pt"))+
  labs(y = "Cumulative Portfolio Return", x = "Date")
```

The graph above shows that high volatility does not give significantly superior performance. In fact, low volatility is vastly superior to high volatility. Worse still, the path to the end is tremondously more volatile, as one would expect. Furthermore, low volatility provides insurance during crashes as can be seen by the performance of high volatility in 2008.

The story is much worse when one examines the drawdowns.

```{r drawdowns, dpi = 600}
# drawdowns
d.xts <- do.call(cbind, lapply(1:ncol(c.xts), function(j){
  cx <- cummax(c.xts[, j])
  dd <- c.xts[, j] / cx
}))

draw_downs <- data.frame(dates = index(d.xts), d.xts)
```

Code chunk above is taken from the calculates the drawdowns of each portfolio.

The code below is used to create underwater chart.

```{r draw_downs_graph,dpi=600, fig.cap="High volatility underwater chart"}
ggplot(draw_downs, aes(x = dates, y = high, group = 1)) +
  geom_ribbon(aes(ymin = min(high), ymax = high), fill="#cedccb", alpha=1) +
  geom_line(color="#69b3a2", size=1) +
  theme_classic()+
  scale_x_date(expand = c(0,0), date_breaks = "1 year", date_labels =  "%y",
               position = "top")+
  scale_y_continuous(expand = c(0,0))+
  labs(x = "", y = "")+
  theme(legend.title = element_blank(),
        legend.key = element_blank(), legend.text = element_text(size = 10),
        legend.position = "bottom",
        legend.margin=margin(c(-25,-25,-5,-5)),
        plot.margin = unit(c(-0.3,0.1,0.2,-0.4), "cm"))
```

So with a high volatility portfolio, one would not have recovered to this day from the 2008 crash. This portfolio has not reached the previous high from 2008 after nearly 12 years, which is remarkable. In addition, the drawdowns are very severe, with the high volatility portfolio being over 60% underwater since the 2008 financial crisis.

```{r low_vol_drawdowns,dpi=600, fig.cap="low volatility underwater chart"}
ggplot(draw_downs, aes(x = dates, y = low, group = 1)) +
  geom_ribbon(aes(ymin = min(low), ymax = low), fill="#cedccb", alpha=1) +
  geom_line(color="#69b3a2", size=1) +
  theme_classic()+
  scale_x_date(expand = c(0,0), date_breaks = "1 year", date_labels =  "%y",
               position = "top")+
  scale_y_continuous(expand = c(0,0))+
  labs(x = "", y = "")+
  theme(legend.title = element_blank(),
        legend.key = element_blank(), legend.text = element_text(size = 10),
        legend.position = "bottom",
        legend.margin=margin(c(-25,-25,-5,-5)),
        plot.margin = unit(c(-0.3,0.1,0.2,-0.4), "cm"))
```

The low volatility portfolio fares much better, with it recovering in 2011 from the 2008 crash. The severity of the drawdrown is dramatically better too, with the the portfolio mostly hovering at only 10% underwater. 